---
title: "CKME 136 Capstone Forest Fires"
output:
  word_document: default
  html_document: default
  df_print: paged
  rows.print: 25
---
1 Load packages.


```{r echo=TRUE}

#Make sure to use libraries below
library(dplyr)
#install.packages("ggplot2")
library(ggplot2)
#install.packages("ggpubr")
library(ggpubr)
#install.packages("tidyr")
library(tidyr)
#install.packages("scales")
library(scales)
#install.packages("wesanderson")
#library(wesanderson)
#install.packages("viridis")  # Install
library("viridis")           # Load
#theme_set(theme_pubclean())
library("ggrepel")
#install.packages("janitor")
library(janitor)
#install.packages("formattable")
#library(formattable)
#install.packages("DT")
#library(DT)
#install.packages("FSelector")
library(FSelector)
#??FSelector

```
2. Read the "fores fire .csv" file from the following website.


```{r echo=TRUE}

fires <-read.csv('CKME 136 Forest Fire Data.csv',header=T)

```

3. Have a look at the data set. View(train.data)

```{r echo=TRUE}

# Fires

head(fires)
tail(fires)
str(fires)

# Check data types of attributes
sapply(fires, class)
sapply(fires, typeof)

```
4. Extract relevant columns.


```{r echo=TRUE}

# Fires

new_fire <- fires[, c("Cause", "Jurisdiction", "Number", "Protection.zone", "Response.category", "Year")]
#new_fire

```

5. Check for missing values.

```{r echo=TRUE}

# Fire

sum(is.na(new_fire$Cause) == TRUE) # 0 Missing values.
length(new_fire$Cause)

sum(is.na(new_fire$Jurisdiction) == TRUE) # 0 Missing values.
length(new_fire$Jurisdiction)

sum(is.na(new_fire$Number) == TRUE) # 8352 initial missing values for "Number" field. 
length(new_fire$Number)

sum(is.na(new_fire$Protection.zone) == TRUE) # 0 Missing values.
length(new_fire$Protection.zone)

sum(is.na(new_fire$Response.category) == TRUE) # 0 Missing values.
length(new_fire$Response.category)

sum(is.na(new_fire$Year) == TRUE) # 0 Missing values.
length(new_fire$Year)

```

6. Only "Number" has missing rows.  Remove all rows with missing values.


```{r echo=TRUE}

# Fires

# Remove remaining records with missing values.
FireClean <- new_fire[complete.cases(new_fire),]

nrow(FireClean) #11519 rows remaining

```

7. Check attributes after missing rows have been removed.


```{r echo=TRUE}

# Fire

attach(FireClean)

head(FireClean)
tail(FireClean)
str(FireClean)
dim(FireClean) # 11519 rows, 6 columns

# Check data types of attributes
sapply(FireClean, class)


levels(FireClean$Cause)
levels(FireClean$Jurisdiction)
levels(FireClean$Protection.zone)
levels(FireClean$Response.category)
levels(FireClean$Year)
summary(FireClean) # Only the "number" attribute maybe usefull with the summary

```

8. Shorten "Jurisdiction" name

```{r echo=TRUE}

FireClean$Juris_Long <- FireClean$Jurisdiction # Duplicate Jurisdiction column

levels(FireClean$Jurisdiction)[levels(FireClean$Jurisdiction) == "British Columbia"] <- "BC"
levels(FireClean$Jurisdiction)[levels(FireClean$Jurisdiction) == "Alberta"] <- "AB"
levels(FireClean$Jurisdiction)[levels(FireClean$Jurisdiction) == "National parks"] <- "NP"
levels(FireClean$Jurisdiction)[levels(FireClean$Jurisdiction) == "Northwest Territories"] <- "NT"
levels(FireClean$Jurisdiction)[levels(FireClean$Jurisdiction) == "Prince Edward Island"] <- "PE"
levels(FireClean$Jurisdiction)[levels(FireClean$Jurisdiction) == "Yukon"] <- "YT"
levels(FireClean$Jurisdiction)[levels(FireClean$Jurisdiction) == "New Brunswick"] <- "NB"
levels(FireClean$Jurisdiction)[levels(FireClean$Jurisdiction) == "Nova Scotia"] <- "NS"
levels(FireClean$Jurisdiction)[levels(FireClean$Jurisdiction) == "Quebec"] <- "QC"
levels(FireClean$Jurisdiction)[levels(FireClean$Jurisdiction) == "Manitoba"] <- "MB"
levels(FireClean$Jurisdiction)[levels(FireClean$Jurisdiction) == "Newfoundland and Labrador"] <- "NL"
levels(FireClean$Jurisdiction)[levels(FireClean$Jurisdiction) == "Ontario"] <- "ON"
levels(FireClean$Jurisdiction)[levels(FireClean$Jurisdiction) == "Saskatchewan"] <- "SK"
levels(FireClean$Jurisdiction)

```
9. create new column: Cause_Grouped  People vs Lightning

```{r echo=TRUE}

FireClean <- FireClean %>%
  mutate(Cause_Grouped = case_when(
    Cause == "Lightning"  ~ "Lightning",
    TRUE                 ~ "People"
    )
  )

```
10. Create new column: Time1   Group Years

```{r echo=TRUE}

FireClean <- FireClean %>%
  mutate(Time1 = case_when(
    Year <= 1995                 ~ "Early 90s",
    Year >= 1996 & Year <= 2000  ~ "Late 90s",
    Year >= 2001 & Year <= 2005  ~ "Early 10s",
    Year >= 2006 & Year <= 2010  ~ "Late 10s",
    Year >= 2011 & Year <= 2015  ~ "Early 20s",
    Year >= 2016                 ~ "Late 20s"
    )
  )

```
11. Create new column: Time2   Group Years

```{r echo=TRUE}

FireClean <- FireClean %>%
  mutate(Time2 = case_when(
    Year >= 1990 & Year <= 1999  ~ "1990s",
    Year >= 2000 & Year <= 2009  ~ "2000s",
    Year >= 2010 & Year <= 2018  ~ "2010s"
    )
  )

```
12. Group Provinces into regions

```{r echo=TRUE}

FireClean <- FireClean %>% 
  mutate(Region = case_when(
    Jurisdiction %in% c("AB", "MB", "SK")           ~ "Prairie Region",
    Jurisdiction %in% c("BC")                       ~ "Pacific Region",
    Jurisdiction %in% c("NP")                       ~ "National Parks",
    Jurisdiction %in% c("NB", "NL", "NS", "PE")     ~ "Atlantic Region",
    Jurisdiction %in% c("ON", "QC")                 ~ "Central Region",
    Jurisdiction %in% c("YT", "NT")                 ~ "North Region"
      )
  )

```

13. Check structure of attributes again.

```{r echo=TRUE}

attach(FireClean)

head(FireClean)
tail(FireClean)
str(FireClean)
dim(FireClean) # 11519 rows, 6 columns

# Check data types of attributes
sapply(FireClean, class)

levels(FireClean$Cause)
levels(FireClean$Jurisdiction)
levels(FireClean$Protection.zone)
levels(FireClean$Response.category)
levels(FireClean$Year)
summary(FireClean) # Only the "number" attribute maybe usefull with the summary
summary(FireClean$Number) 
summary(FireClean$Year) 

```

14. Visulization, Barplot, Boxplot

```{r echo=TRUE, rows.print=30, cols.print=30}

data <- tbl_df(FireClean)

head(data)

########################################################################################################################################
#11 

pivot3 <- data %>%
select(Year, Number, Cause_Grouped)
head(pivot3)


pivot3 <- data %>% #Groups Cause together and sums Number
  select(Cause_Grouped, Number, Year, Jurisdiction) %>% 
  filter(Jurisdiction == "BC") %>% 
  group_by(Year, Cause_Grouped) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE)) 
  

pivot3 %>% 
  spread(Year, sum_Number)

############
#SUMMING COLUMNS AND ROWS
pivot4 <- pivot3 %>% 
  spread(Year, sum_Number)

pivot4

pivot5 <- pivot4 %>%
  adorn_totals("row") %>% 
  adorn_totals("col") 

pivot5

##################################################
# Test 1
  
ks <- function (x) { number_format(accuracy = 1,
                                   scale = 1/1000,
                                   suffix = "k",
                                   big.mark = ",")(x) }


p <- ggplot(data = pivot3, aes(x = Year,
                               y = sum_Number,
                               color = Cause_Grouped)
            ) +
  geom_bar(stat="identity", width = 0.7, fill="white") +
  # geom_text_repel(aes(label=sum_Number), show_guide = F, position=position_dodge(width=0.4),
  #                 vjust= -2.4, hjust = 0.4, size = 3.8, angle = 0)+
  scale_x_continuous(breaks=1990:2018)+
  #stat_summary(fun.y = sum, aes(label = ..y.., group = Year), geom = "text", vjust= -1.5, show_guide = F)+
  xlab("Cause") +
  ylab("Causes of Forest Fires (1990 - 2018)") +
  scale_y_continuous(name="Sum of Forest Fires", labels = ks)+
  coord_cartesian(ylim = c(0, 5000))+
  labs(title = "Causes of Forest Fires in BC (1990 - 2018)")

p
p + theme(
  axis.text.x = element_text(face = "bold", color = "#993333", hjust = 0 ,size = 9, angle = 90))

########################################################################################################################################
#10-B Practice sum column and rows

pivot3 <- data %>%
select(Region, Number, Cause_Grouped)
head(pivot3)

# attach(pivot3)
# detach(pivot3)

pivot3 <- data %>% #Groups Cause together and sums Number
  select(Region, Number, Cause_Grouped) %>% 
  group_by(Cause_Grouped, Region) %>%
  summarize(sum_Number = sum(Number, na.rm = TRUE))
  
pivot3 %>% 
  spread(Cause_Grouped, sum_Number)

############
#SUMMING COLUMNS AND ROWS
pivot4 <- pivot3 %>% 
  spread(Cause_Grouped, sum_Number)

pivot4

pivot5 <- pivot4 %>%
  adorn_totals("row") %>% 
  adorn_totals("col") 

pivot5

##################################################
#CORRELATION USE FOR pivot3 ONLY
CorDataFrame <- pivot3 %>% 
  spread(Cause_Grouped, sum_Number)

CorDataFrame
sapply(CorDataFrame, is.numeric) # Which columns are numeric?

my_num_data <- CorDataFrame[, sapply(CorDataFrame, is.numeric)] # Subset numeric columns
my_num_data


plot(my_num_data) # Works
cor(my_num_data)


##################################################

# Test 1
  
ks <- function (x) { number_format(accuracy = 1,
                                   scale = 1/1000,
                                   suffix = "k",
                                   big.mark = ",")(x) }


p <- ggplot(data = pivot3,
       aes(x = Region, y = sum_Number, color=Cause_Grouped)) +
  geom_bar(position = "dodge", stat="identity", width = 0.7, fill="white") +
  geom_text_repel(aes(label=sum_Number), show_guide = F, position=position_dodge(width=0.4),
                  vjust= -2.4, hjust = 0.4, size = 3.8, angle = 0)+
  xlab("Cause") +
  ylab("Causes of Forest Fires (1990 - 2018)") +
  scale_y_continuous(name="Sum of Forest Fires", labels = ks)+
  coord_cartesian(ylim = c(0, 40000))+
  labs(title = "Causes of Forest Fires (1990 - 2018)")

p
p + theme(
  axis.text.x = element_text(face = "bold", color = "#993333", hjust = 1,size = 9, angle = 45))



########################################################################################################################################
#10 

pivot3 <- data %>%
select(Region, Number, Cause_Grouped)
head(pivot3)

# attach(pivot3)
# detach(pivot3)

pivot3 <- data %>% #Groups Cause together and sums Number
  select(Region, Number, Cause_Grouped) %>% 
  group_by(Cause_Grouped, Region) %>%
  summarize(sum_Number = sum(Number, na.rm = TRUE)) 
  

pivot3 %>% 
  spread(Cause_Grouped, sum_Number)


##################################################
#CORRELATION
CorDataFrame <- pivot3 %>% 
  spread(Cause_Grouped, sum_Number)

CorDataFrame
sapply(CorDataFrame, is.numeric) # Which columns are numeric?

my_num_data <- CorDataFrame[, sapply(CorDataFrame, is.numeric)] # Subset numeric columns
my_num_data


plot(my_num_data) # Works
cor(my_num_data)


##################################################

# Test 1
  
ks <- function (x) { number_format(accuracy = 1,
                                   scale = 1/1000,
                                   suffix = "k",
                                   big.mark = ",")(x) }


p <- ggplot(data = pivot3,
       aes(x = Region, y = sum_Number, color=Cause_Grouped)) +
  geom_bar(position = "dodge", stat="identity", width = 0.7, fill="white") +
  geom_text_repel(aes(label=sum_Number), show_guide = F, position=position_dodge(width=0.4),
                  vjust= -2.4, hjust = 0.4, size = 3.8, angle = 0)+
  xlab("Cause") +
  ylab("Causes of Forest Fires (1990 - 2018)") +
  scale_y_continuous(name="Sum of Forest Fires", labels = ks)+
  coord_cartesian(ylim = c(0, 40000))+
  labs(title = "Causes of Forest Fires (1990 - 2018)")

p
p + theme(
  axis.text.x = element_text(face = "bold", color = "#993333", hjust = 1,size = 9, angle = 45))

########################################################################################################################################
#9

pivot3 <- data %>%
select(Year, Number, Jurisdiction)
head(pivot3)


pivot3 <- data %>% #Groups Cause together and sums Number
  select(Jurisdiction, Number, Year) %>% 
  group_by(Year, Jurisdiction) %>%
  summarize(sum_Number = sum(Number, na.rm = TRUE)) 
  

pivot3 %>% 
  spread(Year, sum_Number)

#Produces matrix with zero
pivot4 <- pivot3 %>% 
  spread(Year, sum_Number)

pivot4

#Remove NAs from matrix, replace with zero
pivot4[is.na(pivot4)] <- 0
#Matrix now has NAs removed, now replaced with zero
pivot4

############
#SUMMING COLUMNS AND ROWS

pivot5 <- pivot4 %>%
  adorn_totals("row") %>% 
  adorn_totals("col") 

pivot5

##################################################
# Test 4
  
ks <- function (x) { number_format(accuracy = 1,
                                   scale = 1/1000,
                                   suffix = "k",
                                   big.mark = ",")(x) }


p <- ggplot(data = pivot3, aes(x = Year,
                               y = sum_Number,
                               color = Jurisdiction)
            ) +
  geom_line(stat="identity", width = 0.7, fill="white", size = 2) +
  # geom_text_repel(aes(label=sum_Number), show_guide = F, position=position_dodge(width=0.4),
  #                 vjust= -2.4, hjust = 0.4, size = 3.8, angle = 0)+
  scale_color_manual(values = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7",
                                "#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442")) +
  scale_x_continuous(breaks=1990:2018)+
  #stat_summary(fun.y = sum, aes(label = ..y.., group = Year), geom = "text", vjust= -1.5, show_guide = F)+
  xlab("Year") +
  ylab("Number of Forest Fires (1990 - 2018)") +
  scale_y_continuous(name="Sum of Forest Fires", labels = ks)+
  coord_cartesian(ylim = c(0, 4500))+
  labs(title = "Number of Forest Fires for each Province (1990 - 2018)")

p
p + theme(
  axis.text.x = element_text(face = "bold", color = "#993333", hjust = 0 ,size = 9, angle = 90))


##################################################
# Test 3
  
ks <- function (x) { number_format(accuracy = 1,
                                   scale = 1/1000,
                                   suffix = "k",
                                   big.mark = ",")(x) }


p <- ggplot(data = pivot3, aes(x = Year,
                               y = sum_Number,
                               color = Jurisdiction)
            ) +
  geom_line(stat="identity", width = 0.7, fill="white", size = 2) +
  # geom_text_repel(aes(label=sum_Number), show_guide = F, position=position_dodge(width=0.4),
  #                 vjust= -2.4, hjust = 0.4, size = 3.8, angle = 0)+
  scale_x_continuous(breaks=1990:2018)+
  #stat_summary(fun.y = sum, aes(label = ..y.., group = Year), geom = "text", vjust= -1.5, show_guide = F)+
  xlab("Year") +
  ylab("Number of Forest Fires (1990 - 2018)") +
  scale_y_continuous(name="Sum of Forest Fires", labels = ks)+
  coord_cartesian(ylim = c(0, 4500))+
  labs(title = "Number of Forest Fires for each Province (1990 - 2018)")

p
p + theme(
  axis.text.x = element_text(face = "bold", color = "#993333", hjust = 0 ,size = 9, angle = 90))


##################################################
# Test 2
  
ks <- function (x) { number_format(accuracy = 1,
                                   scale = 1/1000,
                                   suffix = "k",
                                   big.mark = ",")(x) }


p <- ggplot(data = pivot3, aes(x = Year,
                               y = sum_Number,
                               color = Jurisdiction)
            ) +
  geom_line(stat="identity", width = 0.7, fill="white") +
  # geom_text_repel(aes(label=sum_Number), show_guide = F, position=position_dodge(width=0.4),
  #                 vjust= -2.4, hjust = 0.4, size = 3.8, angle = 0)+
  scale_x_continuous(breaks=1990:2018)+
  #stat_summary(fun.y = sum, aes(label = ..y.., group = Year), geom = "text", vjust= -1.5, show_guide = F)+
  xlab("Year") +
  ylab("Number of Forest Fires (1990 - 2018)") +
  scale_y_continuous(name="Sum of Forest Fires", labels = ks)+
  coord_cartesian(ylim = c(0, 4500))+
  labs(title = "Number of Forest Fires for each Province (1990 - 2018)")

p
p + theme(
  axis.text.x = element_text(face = "bold", color = "#993333", hjust = 0 ,size = 9, angle = 90))


########################################################################################################################################
#8

pivot3 <- data %>%
select(Year, Number, Cause_Grouped)
head(pivot3)


pivot3 <- data %>% #Groups Cause together and sums Number
  select(Cause_Grouped, Number, Year) %>% 
  group_by(Year, Cause_Grouped) %>%
  summarize(sum_Number = sum(Number, na.rm = TRUE)) 
  

pivot3 %>% 
  spread(Year, sum_Number)

##################################################
# Test 8
  
ks <- function (x) { number_format(accuracy = 1,
                                   scale = 1/1000,
                                   suffix = "k",
                                   big.mark = ",")(x) }


p <- ggplot(data = pivot3, aes(x = Year,
                               y = sum_Number,
                               color = Cause_Grouped)
            ) +
  geom_bar(stat="identity", width = 0.7, fill="white") +
  # geom_text_repel(aes(label=sum_Number), show_guide = F, position=position_dodge(width=0.4),
  #                 vjust= -2.4, hjust = 0.4, size = 3.8, angle = 0)+
  scale_x_continuous(breaks=1990:2018)+
  #stat_summary(fun.y = sum, aes(label = ..y.., group = Year), geom = "text", vjust= -1.5, show_guide = F)+
  xlab("Cause") +
  ylab("Causes of Forest Fires (1990 - 2018)") +
  scale_y_continuous(name="Sum of Forest Fires", labels = ks)+
  coord_cartesian(ylim = c(0, 12000))+
  labs(title = "Causes of Forest Fires (1990 - 2018)")

p
p + theme(
  axis.text.x = element_text(face = "bold", color = "#993333", hjust = 0 ,size = 9, angle = 90))


########################################################################################################################################
#7

pivot3 <- data %>%
select(Jurisdiction, Number, Year)
head(pivot3)

pivot3 <- data %>% #Groups Cause together and sums Number
  select(Jurisdiction, Number, Year) %>% 
  group_by(Year, Jurisdiction) %>%
  summarize(sum_Number = sum(Number, na.rm = TRUE)) 
  
#Produces matrix with zero
pivot4 <- pivot3 %>% 
  spread(Year, sum_Number)

pivot4

#Remove NAs from matrix, replace with zero
pivot4[is.na(pivot4)] <- 0
#Matrix now has NAs removed, now replaced with zero
pivot4

############
#SUMMING COLUMNS AND ROWS

pivot5 <- pivot4 %>%
  adorn_totals("row") %>% 
  adorn_totals("col") 

pivot5

##################################################

boxplot(sum_Number~Year,
  data=pivot3,
  main="Distribution of number of fires for all Provinces per year \n(1990 - 2018)",
  xlab="Year",
  ylab="Number of Fires",
  col=c("blue", "indianred1","red","orange","purple","green",
    "pink","brown", "mediumpurple2","olivedrab1"),
  border="black"
)


########################################################################################################################################
#6 Boxplot
pivot3 <- data %>%
select(Jurisdiction, Number, Year)
head(pivot3)

pivot3 <- data %>% #Groups Cause together and sums Number
  select(Jurisdiction, Number, Year) %>% 
  group_by(Year, Jurisdiction) %>%
  summarize(sum_Number = sum(Number, na.rm = TRUE)) 
  

pivot3 %>% 
  spread(Jurisdiction, sum_Number)

#Produces matrix with zero
pivot4 <- pivot3 %>% 
  spread(Jurisdiction, sum_Number)

pivot4

#Remove NAs from matrix, replace with zero
pivot4[is.na(pivot4)] <- 0
#Matrix now has NAs removed, now replaced with zero
pivot4

############
#SUMMING COLUMNS AND ROWS

pivot5 <- pivot4 %>%
  adorn_totals("row") %>% 
  adorn_totals("col") 

pivot5

##################################################


##################################################
#CORRELATION
CorDataFrame <- pivot3 %>% 
  spread(Jurisdiction, sum_Number)


CorDataFrame
sapply(CorDataFrame, is.numeric) # Which columns are numeric?

my_num_data <- CorDataFrame[, sapply(CorDataFrame, is.numeric)] # Subset numeric columns
my_num_data


plot(my_num_data) # Works
cor(my_num_data)

##################################################
# Test 1

boxplot(sum_Number~Jurisdiction,
  data=pivot3,
  main="Distribution of number of fires in each Province per year \n(1990 - 2018)",
  xlab="Provinces",
  ylab="Number of Fires",
  col=c("blue", "indianred1","red","orange","purple","green",
    "pink","brown", "mediumpurple2","olivedrab1"),
  border="black"
)

##################################################

########################################################################################################################################
#5 Boxplot

pivot3 <- data %>%
select(Jurisdiction, Number, Cause_Grouped)
head(pivot3)

pivot3 <- data %>% #Groups Cause together and sums Number
  select(Jurisdiction, Number, Cause_Grouped) %>% 
  group_by(Cause_Grouped, Jurisdiction) %>%
  summarize(sum_Number = sum(Number, na.rm = TRUE)) 
  
pivot3 %>% 
  spread(Cause_Grouped, sum_Number)



##################################################

# Test 2
  

boxplot(sum_Number~Cause_Grouped,
  data=pivot3,
  main="Distribution of number of fires in each Province for each cause \n(1990 - 2018)",
  xlab="Cause",
  ylab="Number of Fires",
  col=c("blue", "indianred1","red","orange","purple","green",
    "pink","brown", "mediumpurple2","olivedrab1"),
  border="black"
)

########################################################################################################################################
#4

pivot3 <- data %>%
select(Jurisdiction, Number, Cause_Grouped)
head(pivot3)

pivot3 <- data %>% #Groups Cause together and sums Number
  select(Jurisdiction, Number, Cause_Grouped) %>% 
  group_by(Cause_Grouped, Jurisdiction) %>%
  summarize(sum_Number = sum(Number, na.rm = TRUE)) 
  
pivot3 %>% 
  spread(Cause_Grouped, sum_Number)


##################################################
# Test 5
  
ks <- function (x) { number_format(accuracy = 1,
                                   scale = 1/1000,
                                   suffix = "k",
                                   big.mark = ",")(x) }


p <- ggplot(data = pivot3, aes(x = Jurisdiction,
                               y = sum_Number,
                               color = Cause_Grouped)
            ) +
  geom_bar(stat="identity", width = 0.7, fill="white") +
  # geom_text_repel(aes(label=sum_Number), show_guide = F, position=position_dodge(width=0.4),
  #                 vjust= -2.4, hjust = 0.4, size = 3.8, angle = 0)+
  stat_summary(fun.y = sum, aes(label = ..y.., group = Jurisdiction), geom = "text", vjust= -1.5, show_guide = F)+
  xlab("Cause") +
  ylab("Causes of Forest Fires (1990 - 2018)") +
  scale_y_continuous(name="Sum of Forest Fires", labels = ks)+
  coord_cartesian(ylim = c(0, 60000))+
  labs(title = "Causes of Forest Fires (1990 - 2018)")

p
p + theme(
  axis.text.x = element_text(face = "bold", color = "#993333", hjust = 1,size = 9, angle = 0))


########################################################################################################################################
#3
pivot3 <- data %>%
select(Jurisdiction, Number, Cause_Grouped)
head(pivot3)

# attach(pivot3)
# detach(pivot3)

pivot3 <- data %>% #Groups Cause together and sums Number
  select(Jurisdiction, Number, Cause_Grouped) %>% 
  group_by(Cause_Grouped, Jurisdiction) %>%
  summarize(sum_Number = sum(Number, na.rm = TRUE)) 
  
pivot3 %>% 
  spread(Cause_Grouped, sum_Number)


##################################################
#CORRELATION
CorDataFrame <- pivot3 %>% 
  spread(Cause_Grouped, sum_Number)

CorDataFrame
sapply(CorDataFrame, is.numeric) # Which columns are numeric?

my_num_data <- CorDataFrame[, sapply(CorDataFrame, is.numeric)] # Subset numeric columns
my_num_data


plot(my_num_data) # Works
cor(my_num_data)


##################################################

# Test 11
  
ks <- function (x) { number_format(accuracy = 1,
                                   scale = 1/1000,
                                   suffix = "k",
                                   big.mark = ",")(x) }


p <- ggplot(data = pivot3,
       aes(x = Jurisdiction, y = sum_Number, color=Cause_Grouped)) +
  geom_bar(position = "dodge", stat="identity", width = 0.7, fill="white") +
  geom_text_repel(aes(label=sum_Number), show_guide = F, position=position_dodge(width=0.4),
                  vjust= -2.4, hjust = 0.4, size = 3.8, angle = 0)+
  xlab("Cause") +
  ylab("Causes of Forest Fires (1990 - 2018)") +
  scale_y_continuous(name="Sum of Forest Fires", labels = ks)+
  coord_cartesian(ylim = c(0, 40000))+
  labs(title = "Causes of Forest Fires (1990 - 2018)")

p
p + theme(
  axis.text.x = element_text(face = "bold", color = "#993333", hjust = 1,size = 9, angle = 0))

########################################################################################################################################
#2
#Below works with percentage totals
GCause_HvsL <- data %>% #Groups Cause together and sums Number
  group_by(Cause_Grouped) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE)) %>% 
  mutate(rel.freq = paste0(round(100 * sum_Number/sum(sum_Number),0),"%"))

GCause_HvsL

##################################################

# Boxplot GCause 11, testing
  
ks <- function (x) { number_format(accuracy = 1,
                                   scale = 1/1000,
                                   suffix = "k",
                                   big.mark = ",")(x) }

p <- ggplot(data = GCause_HvsL,
       aes(x = Cause_Grouped, y = sum_Number, color=Cause_Grouped)) +
  geom_bar(stat="identity", width = 0.3, fill="white") +
  geom_text(aes(label=sum_Number), position=position_dodge(width=0.9), vjust=-0.15)+
  xlab("Cause") +
  ylab("Causes of Forest Fires (1990 - 2018)") +
  scale_y_continuous(name="Sum of Forest Fires", labels = ks)+
  coord_cartesian(ylim = c(0, 125000))+
  labs(title = "Causes of Forest Fires (1990 - 2018)")

p
p + theme(
  axis.text.x = element_text(face = "bold", color = "#993333", hjust = 1,size = 9, angle = 45), legend.position = "none")

########################################################################################################################################
#1
#Below works with percentage totals
GCause <- data %>% #Groups Cause together and sums Number
  group_by(Cause) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE)) %>% 
  mutate(rel.freq = paste0(round(100 * sum_Number/sum(sum_Number),0),"%"))

GCause

##################################################

ks <- function (x) { number_format(accuracy = 1,
                                   scale = 1/1000,
                                   suffix = "k",
                                   big.mark = ",")(x) }

p <- ggplot(data = GCause,
       aes(x = Cause, y = sum_Number, color=Cause)) +
  geom_bar(stat="identity", fill="white") +
  geom_text(aes(label=sum_Number), position=position_dodge(width=0.9), vjust=-0.15)+
  xlab("Cause") +
  ylab("Causes of Forest Fires (1990 - 2018)") +
  scale_y_continuous(name="Sum of Forest Fires", labels = ks)+
  coord_cartesian(ylim = c(0, 105000))+
  labs(title = "Causes of Forest Fires (1990 - 2018)")

p
p + theme(
  axis.text.x = element_text(face = "bold", color = "#993333", hjust = 1,size = 9, angle = 45), legend.position = "none")

########################################################################################################################################

#Sample code for various statistics
data %>% #Groups Cause, Year together and sums Number
  group_by(Cause, Year) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE))    

#Below code works
data %>% #Groups Cause, Year together and sums Number
  group_by(Cause, Jurisdiction, Year) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE))    

#Below code works
data %>% #Groups Cause, Year together and sums Number
  group_by(Jurisdiction, Year) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE))    

#Below code works
data %>% #Groups Cause, Year together and sums Number
  group_by(Year, Jurisdiction) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE))    

#Below code works
data %>% #Groups Cause, Year together and sums Number
  group_by(Jurisdiction) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE))    

#Below code works
data %>% #Groups Cause, Year together and sums Number
  group_by(Jurisdiction) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE))    

#######################################################
#Below code works
data %>% #Groups Cause, Year, Jurisdiction and sums Number
  group_by(Cause, Year, Jurisdiction) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE)) %>% 
  filter(Jurisdiction == "BC") %>%
  filter(Cause == "Lightning")

#Below code works
data %>% #Groups Cause together and sums Number
  group_by(Cause, Year, Jurisdiction) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE)) %>% 
  filter(Jurisdiction == "AB") %>%
  filter(Cause == "Lightning")

#Below works great, show max number of fires and year for Alberta
data %>% #Groups Cause together and sums Number
  group_by(Jurisdiction, Cause, Year) %>% 
  summarize(max_Number = max(Number, na.rm = TRUE)) %>% 
  top_n(n=1) %>% 
  filter(Jurisdiction == "AB") %>%
  filter(Cause == "Lightning") 

#Below works great, Top 10 years for Lightning in Alberta
data %>% #Groups Cause together and sums Number
  group_by(Jurisdiction, Cause, Year) %>% 
  summarize(max_Number = max(Number, na.rm = TRUE)) %>% 
  top_n(n=10) %>% 
  filter(Jurisdiction == "AB") %>%
  filter(Cause == "Lightning") %>% 
  arrange(desc(max_Number))

#Below works great, Top 10 years for Lightning in Alberta
data %>% #Groups Cause together and sums Number
  group_by(Jurisdiction, Cause, Year) %>% 
  summarize(max_Number = max(Number, na.rm = TRUE)) %>% 
  top_n(n=10) %>% 
  filter(Jurisdiction == "AB") %>%
  filter(Cause == "Lightning")
  
#Below works great, Top 10 years for Lightning in Alberta
Alberta <- data %>% #Groups Cause together and sums Number
  group_by(Jurisdiction, Cause, Year) %>% 
  summarize(max_Number = max(Number, na.rm = TRUE)) %>% 
  top_n(n=10) %>% 
  filter(Jurisdiction == "AB") %>%
  filter(Cause == "Lightning")

Alberta

hist(Alberta$Year)
hist(Alberta$max_Number)

#Below works
data %>% #Groups Cause together and sums Number
  group_by(Jurisdiction, Cause, Year) %>% 
  summarize(max_Number = max(Number, na.rm = TRUE)) %>% 
  filter(Jurisdiction == "AB") %>%
  filter(Cause == "Lightning") %>% 
  filter(Year > "2003")
  
#Below works, Top 10 after 2003
data %>% #Groups Cause together and sums Number
  group_by(Jurisdiction, Cause, Year) %>% 
  summarize(max_Number = max(Number, na.rm = TRUE)) %>% 
  filter(Jurisdiction == "AB") %>%
  filter(Cause == "Lightning") %>% 
  filter(Year > "2003") %>% 
  arrange(desc(max_Number))

data %>% 
  group_by(Jurisdiction, Cause, Year) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE))


#Below works, Top 10 after 2003
data %>% #Groups Cause together and sums Number
  group_by(Jurisdiction, Cause, Year) %>% 
  summarize(max_Number = max(Number, na.rm = TRUE)) %>% 
  filter(Jurisdiction == "AB") %>%
  filter(Cause == "Lightning") %>% 
  filter(Year > "2003") %>% 
  arrange(desc(max_Number))

################################################

data %>% #Groups Cause together and sums Number.  Also provides avg for Number of fires.
  group_by(Cause) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE),
    count = n(),
    average_fire = mean(Number, na.rm = TRUE))

data %>% #Groups Cause together and sums Number.  Also provides avg for Number of fires.
  group_by(Cause) %>% 
  summarize(
    sum_Number = sum(Number, na.rm = TRUE),
    count = n(),
    average_fire = mean(Number, na.rm = TRUE),
    total = sum(sum_Number)
    )

data %>% #Groups Year and sums Number
  group_by(Year) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE))

data %>% #Groups Jurisdiction and sums Number
  group_by(Jurisdiction) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE))

data %>% 
  group_by(Year) %>%
  group_by(Jurisdiction) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE))

data %>% 
  group_by(Jurisdiction) %>%
  group_by(Year) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE))

data %>% 
  group_by(Jurisdiction) %>%
  group_by(Year) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE))

data %>% 
  group_by(Jurisdiction) %>%
  group_by(Year) %>% 
  summarize(sum_Number = sum(Number, na.rm = TRUE))

data %>% 
  group_by(Jurisdiction, Cause, Year) %>%
  summarize(sum_Number = sum(Number, na.rm = TRUE))

###################


```

.Misc code to be entered.

```{r echo=TRUE}


```